---
title: "Getting Started with uqcca"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with uqcca}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`uqcca` packages the sparse CCA uncertainty-quantification workflows developed
in this repository so they can be shared with collaborators as a conventional R
package. The package keeps the original `ccar3` fitting backend, then adds
three uncertainty summaries:

1. Bootstrap uncertainty for the sparse CCA loadings and canonical
   correlations.
2. Cross-fitted conditional regression for variables associated with the
   opposite-view canonical score after adjusting for the remaining features.
3. Cross-fitted marginal feature-score regression for cross-loadings or
   structure coefficients.

The package also includes simulation helpers for benchmarking coverage, false
discovery, power, and subspace recovery.

# Installation

Install the package from a local checkout:

```{r}
devtools::install("/path/to/UQ-CCA")
```

The sparse CCA fitting functions expect access to `ccar3`. You can either
install `ccar3` as a package or point `uqcca` to a local source checkout:

```{r}
options(uqcca.ccar3_path = "/path/to/ccar3")
```

or equivalently:

```{r}
Sys.setenv(UQCCA_CCAR3_PATH = "/path/to/ccar3")
```

The simulation helpers additionally use the external `CCAR3_code` repository:

```{r}
options(uqcca.ccar3_code_path = "/path/to/CCAR3_code")
```

# Alcohol Example

The alcohol multi-omics example can be loaded either from the `CVR` package or
from a local `.rda` file:

```{r}
library(uqcca)

alcohol <- load_alcohol_example()
str(alcohol, max.level = 1)
```

Fit sparse CCA with cross-validated `ecca`:

```{r}
fit <- fit_ecca_cv(
  X = alcohol$X,
  Y = alcohol$Y,
  r = 2,
  lambdas = c(0.005, 0.01, 0.02, 0.05, 0.1, 0.2),
  kfolds = 5,
  preprocess_mode = "center"
)
```

# Bootstrap UQ

Bootstrap resampling refits the sparse CCA pipeline, aligns each replicate back
to the reference fit, and summarizes loading uncertainty:

```{r}
boot <- bootstrap_cca_uq(
  X = alcohol$X,
  Y = alcohol$Y,
  reference_fit = fit,
  B = 50,
  seed = 2025,
  refit_mode = "fixed_lambda"
)

gene_boot <- build_bootstrap_loading_table_all(boot, side = "x")
meth_boot <- build_bootstrap_loading_table_all(boot, side = "y")
rho_boot <- build_bootstrap_correlation_table(boot)
subspace_boot <- summarize_bootstrap_subspace(boot, components = 1:2)
```

For collaborator-ready artifacts:

```{r}
write_bootstrap_loading_reports(
  results = list(bootstrap = boot),
  output_dir = "results/alcohol_bootstrap"
)
```

# Cross-Fitted Regression UQ

The legacy Cauchy-combination workflow remains available:

```{r}
crossfit <- crossfit_cca_inference(
  X = alcohol$X,
  Y = alcohol$Y,
  reference_fit = fit,
  K = 5,
  seed = 2026,
  variance_shrinkage = "james-stein"
)
```

The marginal and conditional estimators target different inferential questions:

```{r}
marginal <- crossfit_cca_marginal_inference(
  X = alcohol$X,
  Y = alcohol$Y,
  reference_fit = fit,
  K = 5,
  seed = 2026
)

conditional <- crossfit_cca_conditional_inference(
  X = alcohol$X,
  Y = alcohol$Y,
  reference_fit = fit,
  K = 5,
  seed = 2026,
  method = "auto"
)
```

The high-level wrapper runs the full alcohol workflow in one call:

```{r}
results <- run_alcohol_uq_comparison(
  n_boot = 25,
  K = 5,
  preprocess_mode = "center"
)

print_alcohol_uq_summary(results)
```

# Simulation Benchmark

The package includes reduced-rank regression simulations with known ground
truth. This is useful for reporting coverage, power, false discovery rates, and
subspace recovery:

```{r}
sim <- run_rrr_uq_simulation(
  n_sims = 20,
  n = 500,
  p = 100,
  q = 20,
  r = 2,
  nnzeros = 10,
  n_boot = 25,
  K = 2,
  preprocess_mode = "center"
)

sim$summary$overall
sim$summary$subspace
```

To write the benchmark tables and plots:

```{r}
write_rrr_uq_simulation_reports(
  simulation_results = sim,
  output_dir = "results/simulation_rrr_uq"
)
```

# Installed Scripts

If you want collaborators to run the packaged workflows as standalone scripts,
they are installed under:

```{r}
system.file("scripts", package = "uqcca")
```

Those scripts correspond to:

- `sparse_cca_alcohol.R`
- `simulate_rrr_uq_power.R`
- `simulate_rrr_uq_by_q.R`

They honor the same `UQCCA_*` environment variables documented above.
